### 2021-01-20 MSDS6372Project1.Rmd
Nick's initial rmd update

---
title: "MSDS6372Project1"
author: "Nicole Norelli, Alex Gilbert & Mingyang Nick YU"
date: "1/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Automible MSRP Analysis and Prediction

#### Objecive 1
We are trying to analyse the importance of the "Popularity" variable. While the details of this variable is vague, it was created from social media, and the "higher ups" are curious how much general popularity can play a role in the retail price of a vehicle.

**Data Import and Cleaning...**
```{r dataINC}
library(tidyverse)
# You need to adjust the work environment to run this code
setwd("/Users/mingyang/Desktop/SMU/Applied Statistics/MSDS6372Project1")
# Read in data under variable CarData
CarData = read.csv("data1.csv")
# In case there are NA character values
CarData[CarData=='NA'] <- NA
# Check the reading - can later be deleted
head(CarData)
str(CarData)
dim(CarData)
# Convert factor variables
variables.to.factor = c("Make","Model","Engine.Fuel.Type","Transmission.Type","Driven_Wheels","Market.Category","Vehicle.Size","Vehicle.Style")
CarData[variables.to.factor] = lapply(CarData[variables.to.factor],factor)
# Look at summary of data
summary(CarData)
summary(CarData$Engine.Fuel.Type)

missingData = CarData[rowSums(is.na(CarData)) > 0,]
missingData
dim(missingData)
```

### 2021-01-20 EDA Project 1 NN 6372.Rmd
Add files via upload

---
title: "EDA 6372 Project 1 NN"
author: "Nicole Norelli"
date: "1/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Read in data, fix factors, examine NAs
cars = read.csv('/Users/nicolenorelli/Documents/DS6372/Project1Details_2021/data1.csv')
cars$Make = as.factor(cars$Make)
cars$Transmission.Type = as.factor(cars$Transmission.Type)
cars$Driven_Wheels = as.factor(cars$Driven_Wheels)
cars$Vehicle.Size = as.factor(cars$Vehicle.Size)
cars$Vehicle.Style = as.factor(cars$Vehicle.Style)
cars$Engine.Fuel.Type = as.factor(cars$Engine.Fuel.Type)
cars$Model = as.factor(cars$Model)

# Need a closer look at Market.Category
# Might want to split up the multiple categories
# There are a bunch of NAs in there too as "N/A"

sapply(cars,function(x) sum(is.na(x))) 
# 6 in Number.of.Doors, 30 in Engine.Cylinders, 69 in Engine.HP
# most of the cylinder NAs are the Mazda rx-8, which doesn't really have cylinders

summary(cars)
```

```{r}
library(tidyverse)
# A look at MSRP
cars %>% ggplot(aes(x=MSRP)) + geom_histogram()
cars %>% ggplot(aes(x=log(MSRP))) + geom_histogram()
cars %>% filter(MSRP<1000000) %>% ggplot(aes(x=MSRP)) + geom_histogram()
cars %>% filter(MSRP<1000000) %>% ggplot(aes(x=log(MSRP))) + geom_histogram()

# Summary MSRP stats for most Factors
t(aggregate(MSRP~Engine.Cylinders, data=cars, summary))
t(aggregate(MSRP~Engine.Fuel.Type, data=cars, summary))
t(aggregate(MSRP~Vehicle.Size, data=cars, summary))
t(aggregate(MSRP~Transmission.Type, data=cars, summary))
t(aggregate(MSRP~Vehicle.Style, data=cars, summary))
t(aggregate(MSRP~Driven_Wheels, data=cars, summary))
```

```{r}
# Visualizations of MSRP and possible predictor variables
plot(as.factor(cars$Engine.Cylinders), cars$MSRP, xlab="Cylinders",ylab="MSRP")
plot(as.factor(cars$Engine.Cylinders), log(cars$MSRP), xlab="Cylinders",ylab="logMSRP")
p1 = cars %>% ggplot(aes(x=Engine.Fuel.Type,y=MSRP)) + geom_boxplot()
p1 + theme(axis.text.x = element_text(angle=45, hjust=1))
cars %>% ggplot(aes(x=Transmission.Type,y=MSRP)) + geom_boxplot()
cars %>% ggplot(aes(x=Driven_Wheels,y=MSRP)) + geom_boxplot()
cars %>% ggplot(aes(x=Vehicle.Size,y=MSRP)) + geom_boxplot()
p = cars %>% ggplot(aes(x=Vehicle.Style,y=MSRP)) + geom_boxplot()
p + theme(axis.text.x = element_text(angle=45, hjust=1))
```

```{r}
# MSRP correlations with numeric predictors
library(GGally)
cars %>% select(MSRP, Year, Popularity) %>% ggpairs()
cars %>% select(MSRP, Engine.HP, Engine.Cylinders, highway.MPG, city.mpg) %>% ggpairs()
cars %>% select(MSRP, Year, Popularity) %>% ggpairs()
cars %>% filter (!is.na(Engine.HP)) %>% select(MSRP, Engine.HP, Engine.Cylinders, highway.MPG, city.mpg) %>% ggpairs()
# temporarily removed NAs for convenience in this graph

# Multicollinearity examination
pairs(cars[,c(3,5,6,9,13,14,15,16)])
```

```{r}

```

### 2021-01-20 MSDS6372Project1.Rmd
fixing Engine.Fuel.Type and Mazda RX model missing values

@@
 CarData = read.csv("data1.csv")
-# In case there are NA character values
-CarData[CarData=='NA'] <- NA
+# Fixing three empty values in CarData$Engine.Fuel.Type, after research they are regular unleaded
+CarData$Engine.Fuel.Type = ifelse(CarData$Engine.Fuel.Type=="","regular unleaded",CarData$Engine.Fuel.Type)
+# Fixing Mazda RX model Engine Cylinders to 0
+CarData$Engine.Cylinders = ifelse(CarData$Make=="Mazda"&is.na(CarData$Engine.Cylinders),0,CarData$Engine.Cylinders)
 # Check the reading - can later be deleted
@@
 summary(CarData)
-summary(CarData$Engine.Fuel.Type)
+summary(CarData$Market.Category)
 

### 2021-01-20 EDA Project 1 NN 6372.Rmd
Update EDA Project 1 NN 6372.Rmd
Updated EDA.  Some exploration and commentary on missing values

@@
 summary(cars)
+```
+
+```{r}
+# A closer look at some of the missing values
+cars %>% filter(is.na(Number.of.Doors))
+# Ferrari FF 2013 -  research indicates this has 2 doors
+# Tesla Model S 2016 -  research indicates this has 4 doors
+cars %>% filter(Engine.Fuel.Type=='')
+# Suzuki Verona 2004 - research indicates "regular unleaded"
+cars %>% filter(is.na(Engine.Cylinders))
+# all Mazda RX-8/7 or electric.  Mazda RX-8 & 7 don't have cylinders
+cars %>% filter(Engine.Cylinders==0) # all electric
+cars %>% filter(Engine.Fuel.Type=="electric")
+# there's 66 electric engines.  they all list either 0 or NA for cylinder.  Safe to put 0 I think
+cars %>% filter(is.na(Engine.HP))
+# a wider variety.  impute/remove?
 ```

### 2021-01-20 MSDS6372Project1.Rmd
more imputation

--- 2021-01-20 MSDS6372Project1.Rmd
@@
 CarData = read.csv("data1.csv")
-# In case there are NA character values
-CarData[CarData=='NA'] <- NA
+# Fixing three empty values in CarData$Engine.Fuel.Type, after research they are regular unleaded
+CarData$Engine.Fuel.Type = ifelse(CarData$Engine.Fuel.Type=="","regular unleaded",CarData$Engine.Fuel.Type)
+# Fixing Engine Cylinders to 0 -- Mazda RX or Electrical Vehicle
+CarData$Engine.Cylinders = ifelse(is.na(CarData$Engine.Cylinders),0,CarData$Engine.Cylinders)
+# Fixing Tesla Model S with Number.of.Doors missing
+CarData$Number.of.Doors = ifelse(CarData$Make=="Tesla"&is.na(CarData$Number.of.Doors),4,CarData$Number.of.Doors)
+# Fixing Tesla HP with mean HP found online of 510
+#CarData$Number.of.Doors = ifelse(CarData$Make=="Tesla"&is.na(CarData$Number.of.Doors),4,CarData$Number.of.Doors)
+# Fixing Ferrari FF 2013 Number.of.Doors to 2 -- Google research
+CarData$Number.of.Doors = ifelse(CarData$Make=="Ferrari"&is.na(CarData$Number.of.Doors),2,CarData$Number.of.Doors)
 # Check the reading - can later be deleted
@@
 dim(missingData)
+
+CarData %>% filter(Make=="Tesla")
 ```

### 2021-01-21 MSDS6372Project1.Rmd
adding program to convert Market.Category into column

--- 2021-01-20 MSDS6372Project1.Rmd
@@
 CarData = read.csv("data1.csv")
-# In case there are NA character values
-CarData[CarData=='NA'] <- NA
+# Fixing three empty values in CarData$Engine.Fuel.Type, after research they are regular unleaded
+CarData$Engine.Fuel.Type = ifelse(CarData$Engine.Fuel.Type=="","regular unleaded",CarData$Engine.Fuel.Type)
+# Fixing Engine Cylinders to 0 -- Mazda RX or Electrical Vehicle
+CarData$Engine.Cylinders = ifelse(is.na(CarData$Engine.Cylinders),0,CarData$Engine.Cylinders)
+# Fixing Tesla Model S with Number.of.Doors missing
+CarData$Number.of.Doors = ifelse(CarData$Make=="Tesla"&is.na(CarData$Number.of.Doors),4,CarData$Number.of.Doors)
+# Fixing Tesla HP with mean HP found online of 510
+#CarData$Number.of.Doors = ifelse(CarData$Make=="Tesla"&is.na(CarData$Number.of.Doors),4,CarData$Number.of.Doors)
+# Fixing Ferrari FF 2013 Number.of.Doors to 2 -- Google research
+CarData$Number.of.Doors = ifelse(CarData$Make=="Ferrari"&is.na(CarData$Number.of.Doors),2,CarData$Number.of.Doors)
+
+#Market.Category: Crossover,Diesel,Exotic,Luxury,High-Performance,Factory Tuner,Performance,Flex Fuel,Hatchback,Hybrid,N/A
+marketCat = c("Crossover","Diesel","Exotic","Luxury","High-Performance","Factory Tuner","Performance","Flex Fuel","Hatchback","Hybrid","N/A")
+marketCat = data.frame(marketCat)
+str = "Hatchback,Factory Tuner,High-Performance,N/A";
+grepl(marketCat[11,1],str)
+
+# Add each Market Category as column
+CarData['Crossover'] <- NA
+CarData['Diesel'] <- NA
+CarData['Exotic'] <- NA
+CarData['Luxury'] <- NA
+CarData['High-Performance'] <- NA
+CarData['Factory Tuner'] <- NA
+CarData['Performance'] <- NA
+CarData['Flex Fuel'] <- NA
+CarData['Hatchback'] <- NA
+CarData['Hybrid'] <- NA
+CarData['N/A'] <- NA
+
+str(CarData)
+# Use For loop to fill in Yes/No to each Market Category column
+for(j in 1:dim(marketCat)[1]){
+  CarData[marketCat[j,1]] = ifelse(grepl(marketCat[j,1],CarData$Market.Category),"Yes","No")
+}
+
 # Check the reading - can later be deleted
@@
 # Convert factor variables
-variables.to.factor = c("Make","Model","Engine.Fuel.Type","Transmission.Type","Driven_Wheels","Market.Category","Vehicle.Size","Vehicle.Style")
+variables.to.factor = c("Make","Model","Engine.Fuel.Type","Transmission.Type","Driven_Wheels","Market.Category","Vehicle.Size","Vehicle.Style","Crossover","Diesel","Exotic","Luxury","High-Performance","Factory Tuner","Performance","Flex Fuel","Hatchback","Hybrid","N/A")
 CarData[variables.to.factor] = lapply(CarData[variables.to.factor],factor)
@@
 dim(missingData)
+
+#Looking at Tesla
+#CarData %>% filter(Make=="Tesla")
+
+# Looking at what to do with Market.Category variable
+marketCat = CarData %>% select(MSRP,Market.Category)
+dim(marketCat)
+head(marketCat,100)
+
 ```

### 2021-01-21 MSDS6372Project1.Rmd
update to Performance and High-Performance issue

@@
 
-#Market.Category: Crossover,Diesel,Exotic,Luxury,High-Performance,Factory Tuner,Performance,Flex Fuel,Hatchback,Hybrid,N/A
-marketCat = c("Crossover","Diesel","Exotic","Luxury","High-Performance","Factory Tuner","Performance","Flex Fuel","Hatchback","Hybrid","N/A")
+#Market.Category: Crossover,Diesel,Exotic,Luxury,High-Performance,Factory Tuner,Performance,Flex Fuel,Hatchback,Hybrid,N/A (Note: Performance column will be handled individually becuase of presence of High-Performance)
+marketCat = c("Crossover","Diesel","Exotic","Luxury","High-Performance","Factory Tuner","Flex Fuel","Hatchback","Hybrid","N/A")
 marketCat = data.frame(marketCat)
@@
 }
+# Performance is being handled individually...
+CarData$Performance = ifelse(grepl("Performance",CarData$Market.Category)&(!grepl("High-Performance",CarData$Market.Category)),"Yes","No")
 
+CarData %>% select(Market.Category,Performance,'High-Performance','Flex Fuel')
 # Check the reading - can later be deleted

### 2021-01-21 MSDS6372Project1.Rmd
fixing variable column names

@@
 marketCat = data.frame(marketCat)
-str = "Hatchback,Factory Tuner,High-Performance,N/A";
-grepl(marketCat[11,1],str)
 
@@
 CarData %>% select(Market.Category,Performance,'High-Performance','Flex Fuel')
-# Check the reading - can later be deleted
-head(CarData)
-str(CarData)
-dim(CarData)
+
 # Convert factor variables
@@
 CarData[variables.to.factor] = lapply(CarData[variables.to.factor],factor)
+
+# Adjust column names - get ride of - and space in the column names to make analysis easier
+# Also make column names more uniform
+CarData = CarData %>%
+  rename(
+    EngineFuelType = Engine.Fuel.Type,
+    EngineHP = Engine.HP,
+    EngineCylinders = Engine.Cylinders,
+    TransmissionType = Transmission.Type,
+    DrivenWheels = Driven_Wheels,
+    NumberOfDoors = Number.of.Doors,
+    MarketCategory = Market.Category,
+    VehicleSize = Vehicle.Size,
+    VehicleStyle = Vehicle.Style,
+    HighwayMPG = highway.MPG,
+    CityMPG = city.mpg,
+    HighPerformance = `High-Performance`,
+    FactoryTuner = `Factory Tuner`,
+    FlexFuel = `Flex Fuel`,
+    N_A = `N/A`
+  )
+
 # Look at summary of data
 summary(CarData)
-summary(CarData$Market.Category)
 

### 2021-01-22 MSDS6372Project1.Rmd
some more
forgot to update rmd

@@
 CarData$Number.of.Doors = ifelse(CarData$Make=="Ferrari"&is.na(CarData$Number.of.Doors),2,CarData$Number.of.Doors)
+# Fixing FIAT 500e HP NA value based on research on mean Horsepower
+CarData$Engine.HP = ifelse(CarData$Make=="FIAT"&is.na(CarData$Engine.HP),111,CarData$Engine.HP)
+# Fixing Lincoln Continental HP NA value based on research on mean Horsepower
+CarData$Engine.HP = ifelse(CarData$Make=="Lincoln"&is.na(CarData$Engine.HP),345,CarData$Engine.HP)
+# Fixing Ford Escape HP NA value based on research on mean Horsepower
+CarData$Engine.HP = ifelse(CarData$Make=="Ford"&is.na(CarData$Engine.HP),210,CarData$Engine.HP)
+# Fixing Honda Fit EV HP NA value based on research on mean Horsepower
+CarData$Engine.HP = ifelse(CarData$Make=="Honda"&is.na(CarData$Engine.HP),123,CarData$Engine.HP)
+# Fixing Mitsubishi i-MiEV HP NA value based on research on mean Horsepower
+CarData$Engine.HP = ifelse(CarData$Make=="Mitsubishi"&is.na(CarData$Engine.HP),66,CarData$Engine.HP)
+# Fixing Chevrolet  Impala HP NA value based on research on mean Horsepower
+CarData$Engine.HP = ifelse(CarData$Make=="Chevrolet"&is.na(CarData$Engine.HP),305,CarData$Engine.HP)
+# Fixing Nissan Leaf HP NA value based on research on mean Horsepower
+CarData$Engine.HP = ifelse(CarData$Make=="Nissan"&is.na(CarData$Engine.HP),177,CarData$Engine.HP)
+# Fixing Tesla Model S HP NA value based on research on mean Horsepower
+CarData$Engine.HP = ifelse(CarData$Make=="Tesla"&is.na(CarData$Engine.HP),503,CarData$Engine.HP)
+# Fixing Mercedes-Benz M-Class S HP NA value based on research on mean Horsepower
+CarData$Engine.HP = ifelse(CarData$Make=="Mercedes-Benz"&is.na(CarData$Engine.HP),350,CarData$Engine.HP) 
+# Fixing Toyota RAV4 HP NA value based on research on mean Horsepower
+CarData$Engine.HP = ifelse(CarData$Make=="Toyota"&is.na(CarData$Engine.HP),203,CarData$Engine.HP) 
+# Fixing Kia Soul EV HP NA value based on research on mean Horsepower
+CarData$Engine.HP = ifelse(CarData$Make=="Kia"&is.na(CarData$Engine.HP),109,CarData$Engine.HP)
 
@@
 CarData$Performance = ifelse(grepl("Performance",CarData$Market.Category)&(!grepl("High-Performance",CarData$Market.Category)),"Yes","No")
-
-CarData %>% select(Market.Category,Performance,'High-Performance','Flex Fuel')
 
@@
 # Look at summary of data
-summary(CarData)
-
-missingData = CarData[rowSums(is.na(CarData)) > 0,]
-missingData
-dim(missingData)
-
-#Looking at Tesla
-#CarData %>% filter(Make=="Tesla")
-
-# Looking at what to do with Market.Category variable
-marketCat = CarData %>% select(MSRP,Market.Category)
-dim(marketCat)
-head(marketCat,100)
+summary(CarData$VehicleStyle)
+# Look at missing Data if any
+# missingData = CarData[rowSums(is.na(CarData)) > 0,]
+# missingData
+# dim(missingData)
+write.csv(CarData,"CleanedCarData.csv",row.names = FALSE)
 

### 2021-01-23 MSDS6372Project1.Rmd
update with Forward & Backward methods. Problem with test validation levels...

@@
 ## Automible MSRP Analysis and Prediction
-
-#### Objecive 1
-We are trying to analyse the importance of the "Popularity" variable.
-While the details of this variable is vague, it was created from social media, and the "higher ups" are curious how much general popularity can play a role in the retail price of a vehicle.
 
@@
 ```
+
+#### Objecive 1
+We are trying to analyse the importance of the "Popularity" variable.
+While the details of this variable is vague, it was created from social media, and the "higher ups" are curious how much general popularity can play a role in the retail price of a vehicle.
+
+```{r}
+library(tidyverse)
+library(ggplot2)
+library(GGally)
+# You need to adjust the work environment to run this code
+setwd("/Users/mingyang/Desktop/SMU/Applied Statistics/MSDS6372Project1")
+# Directly import cleaned dataset and start from here
+CarData = read.csv("CleanedCarData.csv")
+# Convert factor variables
+variables.to.factor = c("Make","Model","EngineFuelType","TransmissionType","DrivenWheels","MarketCategory","VehicleSize","VehicleStyle","Crossover","Diesel","Exotic","Luxury","HighPerformance","FactoryTuner","Performance","FlexFuel","Hatchback","Hybrid","N_A")
+CarData[variables.to.factor] = lapply(CarData[variables.to.factor],factor)
+summary(CarData)
+# Plot agsint numeric variables
+CarData %>% select(MSRP,Year,EngineHP,EngineCylinders,NumberOfDoors,HighwayMPG,CityMPG,Popularity) %>% ggpairs()
+# Try Log transformation on MSRP
+CarData$LogMSRP = log(CarData$MSRP)
+# Recheck the logMSRP against numeric variables
+# LogMSRP seem to have improved linear relationship between Year and EngineHP correlations
+CarData %>% select(LogMSRP,Year,EngineHP,EngineCylinders,NumberOfDoors,HighwayMPG,CityMPG,Popularity) %>% ggpairs()
+# Look at Popularity
+boxplot(CarData$Popularity)
+CarData %>% select(LogMSRP,Popularity)%>% ggpairs()
+#Try LogPopularity
+CarData$LogPopularity = log(CarData$Popularity)
+CarData %>% select(LogMSRP,LogPopularity)%>% ggpairs()
+# Look at original MSRP vs Popularity
+CarData %>% select(MSRP,Popularity)%>% ggpairs()
+#LogMSRP and LogPopularity seem to have improved the linearity relationship between the two
+# Look at histogram of LogMSRP
+CarData %>% ggplot(aes(x=LogMSRP))+geom_histogram(bins=100)
+# Drop MSRP and Popularity from the original Dataset
+CarData<-CarData[,-c(15,16)]
+# Drop MarketCategory - 72 levels are not useful for prediction
+CarData <-CarData[,-10]
+str(CarData)
+#Split data into 80% Train 10% Test 10% Validation set
+set.seed(111)
+train.test.index<-sample(1:dim(CarData)[1],10723,replace=F)
+train.test = CarData[train.test.index,]
+validate = CarData[-train.test.index,]
+set.seed(112)
+train.index = sample(1:dim(train.test)[1],9531,replace=F)
+train = train.test[train.index,]
+test = train.test[-train.index,]
+
+# Use Forward variable selection method to select possible variables as predictors
+library(leaps)
+reg.fwd=regsubsets(LogMSRP~.,data=train,method="forward",nvmax=955)
+par(mfrow=c(1,3))
+bics<-summary(reg.fwd)$bic
+plot(1:956,bics,type="l",ylab="BIC",xlab="# of predictors")
+index<-which(bics==min(bics))
+points(index,bics[index],col="red",pch=10)
+
+adjr2<-summary(reg.fwd)$adjr2
+plot(1:956,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
+index<-which(adjr2==max(adjr2))
+points(index,adjr2[index],col="red",pch=10)
+
+rss<-summary(reg.fwd)$rss
+plot(1:956,rss,type="l",ylab="train RSS",xlab="# of predictors")
+index<-which(rss==min(rss))
+points(index,rss[index],col="red",pch=10)
+
+coef(reg.fwd,676)
+# By using Forward method, 676 predictors seem to generate the smallest BIC number
+# This model has selected predictors below:
+# Make+Model+EngineFuelType+TransmissionType+VehicleSize+VehicleStyle+HighwayMPG+CityMPG+Luxury+FactoryTuner+Performance+FlexFuel+Hybrid+N_A+Exotic+Hatchback+LogPopularity
+forward.regression = lm(LogMSRP~Make+Model+EngineFuelType+TransmissionType+VehicleSize+VehicleStyle+HighwayMPG+CityMPG+Luxury+FactoryTuner+Performance+FlexFuel+Hybrid+N_A+Exotic+Hatchback+LogPopularity,data=train)
+summary(forward.regression)
+par(mfrow=c(2,2))
+plot(forward.regression)
+library(olsrr)
+ols_plot_cooksd_bar(forward.regression)
+ols_plot_resid_stand(forward.regression)
+# After observing the Cook's D, there is observation 4659, 9435 has extremely large Cook's D
+# Observation 1523 also has large Cook's D
+train[4659,] # Toyota Yaris iA
+exp(train[4659,25]) 
+train%>% filter(Make=="Toyota") %>% select(Make,Model,Year,LogMSRP)
+# Can't really make sense to eliminate observation 4659
+
+train[9435,] # Ferrari Superamerica 2005
+exp(train[9435,25]) # Exotic Car
+
+#Look at other Exotic Cars
+train%>%filter(Exotic=="Yes")%>% select(Make,Model,Year,LogMSRP,LogPopularity)
+# Don't really make sense to delete observation 4659 either
+
+# Since there are so many categorical variables, and log transformation improvement on Year and EngineHP didn't make it to the final model
+# Next section we will try reimport the dataset without transformation
+```
+
+#### Objective 1 continue without Log transformation - since there are so many categorical variables
+
+```{r}
+library(tidyverse)
+library(ggplot2)
+library(GGally)
+# You need to adjust the work environment to run this code
+setwd("/Users/mingyang/Desktop/SMU/Applied Statistics/MSDS6372Project1")
+# Directly import cleaned dataset and start from here
+CarData = read.csv("CleanedCarData.csv")
+# Convert factor variables
+variables.to.factor = c("Make","Model","EngineFuelType","TransmissionType","DrivenWheels","MarketCategory","VehicleSize","VehicleStyle","Crossover","Diesel","Exotic","Luxury","HighPerformance","FactoryTuner","Performance","FlexFuel","Hatchback","Hybrid","N_A")
+CarData[variables.to.factor] = lapply(CarData[variables.to.factor],factor)
+summary(CarData)
+# Plot agsint numeric variables
+CarData %>% select(MSRP,Year,EngineHP,EngineCylinders,NumberOfDoors,HighwayMPG,CityMPG,Popularity) %>% ggpairs()
+CarData = CarData[,-10] #delete MarketCategory
+# Try different way of sampling by Models
+
+#Without transformation, split dataset
+set.seed(111)
+train.test.index<-sample(1:dim(CarData)[1],10723,replace=F)
+train.test = CarData[train.test.index,]
+validate = CarData[-train.test.index,]
+set.seed(112)
+train.index = sample(1:dim(train.test)[1],9531,replace=F)
+train = train.test[train.index,]
+test = train.test[-train.index,]
+
+str(train)
+
+# Use Forward variable selection method to select possible variables as predictors
+library(leaps)
+reg.fwd=regsubsets(MSRP~.,data=train,method="forward",nvmax=955)
+par(mfrow=c(1,3))
+bics<-summary(reg.fwd)$bic
+plot(1:956,bics,type="l",ylab="BIC",xlab="# of predictors")
+index<-which(bics==min(bics))
+points(index,bics[index],col="red",pch=10)
+
+adjr2<-summary(reg.fwd)$adjr2
+plot(1:956,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
+index<-which(adjr2==max(adjr2))
+points(index,adjr2[index],col="red",pch=10)
+
+rss<-summary(reg.fwd)$rss
+plot(1:956,rss,type="l",ylab="train RSS",xlab="# of predictors")
+index<-which(rss==min(rss))
+points(index,rss[index],col="red",pch=10)
+
+# how many predictors gives the best BIC?
+which(bics==min(bics))
+coef(reg.fwd,345)
+# By looking at this model, it has selected following predictors:
+# Make+Model+EngineFuelType+EngineCylinders+DrivenWheels+VehicleStyle+HighwayMPG+CityMPG+FactoryTuner+Performance+Hybrid+N_A+Exotic+Hatchback
+forward.regression = lm(MSRP~Make+Model+EngineFuelType+EngineCylinders+DrivenWheels+VehicleStyle+HighwayMPG+CityMPG+FactoryTuner+Performance+Hybrid+N_A+Exotic+Hatchback,data=train)
+summary(forward.regression)
+# This model has higher adjusted R-squared compared to the Log transformed model
+par(mfrow=c(2,2))
+plot(forward.regression)
+library(olsrr)
+ols_plot_cooksd_bar(forward.regression)
+ols_plot_resid_stand(forward.regression)
+# Cook's D plot looks okay - Much better than Log Transformed Plot
+#Observation 577, 230 and 7329 seem to have problem on the residual plot
+train[577,]
+train[230,]
+train[7329,]
+
+########################################
+# Try Backward elimination model.#######
+########################################
+reg.bwd=regsubsets(MSRP~.,data=train,method="backward",nvmax=955)
+par(mfrow=c(1,3))
+bics<-summary(reg.bwd)$bic
+plot(1:956,bics,type="l",ylab="BIC",xlab="# of predictors")
+index<-which(bics==min(bics))
+points(index,bics[index],col="red",pch=10)
+
+adjr2<-summary(reg.bwd)$adjr2
+plot(1:956,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
+index<-which(adjr2==max(adjr2))
+points(index,adjr2[index],col="red",pch=10)
+
+rss<-summary(reg.bwd)$rss
+plot(1:956,rss,type="l",ylab="train RSS",xlab="# of predictors")
+index<-which(rss==min(rss))
+points(index,rss[index],col="red",pch=10)
+
+# how many predictors gives the best BIC?
+which(bics==min(bics))
+coef(reg.bwd,307)
+# Predictor for backward elimination method based on lowest BIC value
+# Make+Model+EngineFuelType+EngineCylinders+DrivenWheels+VehicleStyle+HighwayMPG+Performance+Exotic+Hatchback
+backward.regression = lm(MSRP~Make+Model+EngineFuelType+EngineCylinders+DrivenWheels+VehicleStyle+HighwayMPG+Performance+Exotic+Hatchback,data=train)
+summary(backward.regression)
+# Forward select has higher adjusted R-squared compared to backward method
+par(mfrow=c(2,2))
+plot(backward.regression)
+
+ols_plot_cooksd_bar(backward.regression)
+ols_plot_resid_stand(backward.regression)
+#Observation 577, 230 and 7329 seem to have problem on the residual plot - Same as forward selection
+
+############################################
+# Compare Forward vs. Backward on Test set #
+############################################
+# Test result
+test.result = test$MSRP
+str(test)
+test = test[,-15] #delete MSRP for prediction
+#Forward
+forwardPred = predict(forward.regression,test)
+
+```

### 2021-01-25 MSDS6372Project1.Rmd
fixing train/validate/test set split: line 231 to line 253

@@
 
-#Without transformation, split dataset
+######### Try Something else to split data ############
+# First 842 are Models that only have 5 observations or less
+ModelList = CarData%>% group_by(Model) %>% mutate(ModelCount = n()) %>% arrange(ModelCount)
+#arrange(ModelList,desc(count))%>%print(n = Inf)
+# Can only split train/test/validate set on Models that has 6 observations or more
+ModelList = ModelList[,-27]
+Not.Able.To.Split = ModelList[(1:842),]
+Able.To.Split = ModelList[-(1:842),]
+
+# Split Train, test, validate based on able to Train Data...
 set.seed(111)
-train.test.index<-sample(1:dim(CarData)[1],10723,replace=F)
-train.test = CarData[train.test.index,]
-validate = CarData[-train.test.index,]
+validate.set.index<-sample(1:dim(Able.To.Split)[1],1191,replace=F)
+validate = Able.To.Split[validate.set.index,]
+remains1 = Able.To.Split[-validate.set.index,]
 set.seed(112)
-train.index = sample(1:dim(train.test)[1],9531,replace=F)
-train = train.test[train.index,]
-test = train.test[-train.index,]
+test.index = sample(1:dim(remains1)[1],1191,replace=F)
+test = remains1[test.index,]
+train = remains1[-test.index,]
+
+# Joining train set with dataset that been left out
+train = bind_rows(train,Not.Able.To.Split)
+dim(train)
 
 str(train)
-
+str(CarData)
 # Use Forward variable selection method to select possible variables as predictors

### 2021-01-25 MSDS6372Project1.Rmd
forgot to save rmd

@@
   )
+# Look at HighwayMPG outlier identified by Nicole
+CarData %>% filter(HighwayMPG==354)
+# Change observation HighwayMPG 354 to 34
+CarData$HighwayMPG = ifelse(CarData$HighwayMPG==354,34,CarData$HighwayMPG)
 
@@
 summary(CarData$VehicleStyle)
+str(CarData)
 # Look at missing Data if any

### 2021-01-25 MSDS6372Project1.Rmd
update with LASSO

@@
 # Plot agsint numeric variables
-CarData %>% select(MSRP,Year,EngineHP,EngineCylinders,NumberOfDoors,HighwayMPG,CityMPG,Popularity) %>% ggpairs()
+# CarData %>% select(MSRP,Year,EngineHP,EngineCylinders,NumberOfDoors,HighwayMPG,CityMPG,Popularity) %>% ggpairs()
 CarData = CarData[,-10] #delete MarketCategory
@@
 # Use Forward variable selection method to select possible variables as predictors
+########################################
+# Try Forward selection model.##########
+########################################
 library(leaps)
-reg.fwd=regsubsets(MSRP~.,data=train,method="forward",nvmax=955)
+reg.fwd=regsubsets(MSRP~.,data=train,method="forward",nvmax=966)
 par(mfrow=c(1,3))
 bics<-summary(reg.fwd)$bic
-plot(1:956,bics,type="l",ylab="BIC",xlab="# of predictors")
+plot(1:967,bics,type="l",ylab="BIC",xlab="# of predictors")
 index<-which(bics==min(bics))
@@
 adjr2<-summary(reg.fwd)$adjr2
-plot(1:956,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
+plot(1:967,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
 index<-which(adjr2==max(adjr2))
@@
 rss<-summary(reg.fwd)$rss
-plot(1:956,rss,type="l",ylab="train RSS",xlab="# of predictors")
+plot(1:967,rss,type="l",ylab="train RSS",xlab="# of predictors")
 index<-which(rss==min(rss))
@@
 which(bics==min(bics))
-coef(reg.fwd,345)
+coef(reg.fwd,305)
 # By looking at this model, it has selected following predictors:
-# Make+Model+EngineFuelType+EngineCylinders+DrivenWheels+VehicleStyle+HighwayMPG+CityMPG+FactoryTuner+Performance+Hybrid+N_A+Exotic+Hatchback
-forward.regression = lm(MSRP~Make+Model+EngineFuelType+EngineCylinders+DrivenWheels+VehicleStyle+HighwayMPG+CityMPG+FactoryTuner+Performance+Hybrid+N_A+Exotic+Hatchback,data=train)
+# Make+Model+EngineFuelType+VehicleSize+VehicleStyle+Crossover+HighPerformance+TransmissionType+Exotic+Hatchback
+forward.regression = lm(MSRP~Make+Model+EngineFuelType+VehicleSize+VehicleStyle+Crossover+HighPerformance+TransmissionType+Exotic+Hatchback,data=train)
 summary(forward.regression)
@@
 # Cook's D plot looks okay - Much better than Log Transformed Plot
-#Observation 577, 230 and 7329 seem to have problem on the residual plot
-train[577,]
-train[230,]
-train[7329,]
+#Observation 3748,9152,9153 seem to have problem on the residual plot
+summary(train[3748,]) # Ford Focus ... Not sure if we can delete this...
+summary(train[9152,]) # Bugatti ... Delete?
+summary(train[9153,]) # Bugatti ... Delete?
+summary(train[1043,]) # Volvo ... Not seem to cause too much trouble on Cook's D
 
@@
 ########################################
-reg.bwd=regsubsets(MSRP~.,data=train,method="backward",nvmax=955)
+reg.bwd=regsubsets(MSRP~.,data=train,method="backward",nvmax=966)
 par(mfrow=c(1,3))
 bics<-summary(reg.bwd)$bic
-plot(1:956,bics,type="l",ylab="BIC",xlab="# of predictors")
+plot(1:967,bics,type="l",ylab="BIC",xlab="# of predictors")
 index<-which(bics==min(bics))
@@
 adjr2<-summary(reg.bwd)$adjr2
-plot(1:956,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
+plot(1:967,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
 index<-which(adjr2==max(adjr2))
@@
 rss<-summary(reg.bwd)$rss
-plot(1:956,rss,type="l",ylab="train RSS",xlab="# of predictors")
+plot(1:967,rss,type="l",ylab="train RSS",xlab="# of predictors")
 index<-which(rss==min(rss))
@@
 which(bics==min(bics))
-coef(reg.bwd,307)
+coef(reg.bwd,248)
 # Predictor for backward elimination method based on lowest BIC value
-# Make+Model+EngineFuelType+EngineCylinders+DrivenWheels+VehicleStyle+HighwayMPG+Performance+Exotic+Hatchback
-backward.regression = lm(MSRP~Make+Model+EngineFuelType+EngineCylinders+DrivenWheels+VehicleStyle+HighwayMPG+Performance+Exotic+Hatchback,data=train)
+# Make+Model+EngineFuelType+VehicleSize+Crossover+HighPerformance+Hybrid+VehicleStyle+Exotic
+backward.regression = lm(MSRP~Make+Model+EngineFuelType+VehicleSize+Crossover+HighPerformance+Hybrid+VehicleStyle+Exotic,data=train)
 summary(backward.regression)
@@
 ols_plot_resid_stand(backward.regression)
-#Observation 577, 230 and 7329 seem to have problem on the residual plot - Same as forward selection
+#Observation 3748, 9152 and 9153 seem to have problem on the residual plot - Same as forward selection
 
-############################################
-# Compare Forward vs. Backward on Test set #
-############################################
+########################################
+# Try LASSO variable selection #########
+########################################
+library(glmnet)
+#Dummy code categorical predictor variables
+x <- model.matrix(MSRP~.,train)[,-1]
+y <- train$MSRP
+cv.out=cv.glmnet(x,y,alpha=1) #alpha=1 performs LASSO
+par(mfrow=c(1,1))
+plot(cv.out)
+# Look at the coefficient for best lambda
+# coef(cv.out, cv.out$lambda.min)
+# Compute the final lasso model
+lasso.model = glmnet(x,y,alpha = 1,lambda = cv.out$lambda.min)
+
+#Dummy code test set to see performance compare to other methods
+xtest<-model.matrix(MSRP~.,test)[,-1]
+ytest<-test$MSRP #36461071
+
+# See prediction
+lasso.pred=predict(lasso.model ,newx=xtest)
+# See test ASE
+testMSE_LASSO<-mean((ytest-lasso.pred)^2)
+testMSE_LASSO
+
+#####################################################
+# Compare Forward vs. Backward on Test set to LASSO #
+#####################################################
 # Test result
 test.result = test$MSRP
-str(test)
+# str(test)
 test = test[,-15] #delete MSRP for prediction
-#Forward
-forwardPred = predict(forward.regression,test)
+#Forward test results
+forward.pred = predict(forward.regression,test)
+# See forward test ASE
+testMSE_Forward<-mean((test.result-forward.pred)^2)
+testMSE_Forward #46312425
+
+#Backward test results
+backward.pred = predict(backward.regression,test)
+# See backward test ASE
+testMSE_Backward<-mean((test.result-backward.pred)^2)
+testMSE_Backward #48317165
 

### 2021-01-25 MSDS6372Project1.Rmd
update with LASSO seed

@@
 y <- train$MSRP
+set.seed(123) # Need a seed number to keep same results for LASSO
 cv.out=cv.glmnet(x,y,alpha=1) #alpha=1 performs LASSO
@@
 # Look at the coefficient for best lambda
-# coef(cv.out, cv.out$lambda.min)
+coefficients = coef(cv.out, cv.out$lambda.min)
+head(coefficients)
 # Compute the final lasso model
 lasso.model = glmnet(x,y,alpha = 1,lambda = cv.out$lambda.min)
+# Check assumptions
 

### 2021-01-26 MSDS6372Project1.Rmd
deleted MSRP over 1mil updated models accordingly...

@@
 CarData = read.csv("CleanedCarData.csv")
+###################################################################################
+# Restrict prediction range to over cars under 1 million... To improve precision###
+###################################################################################
+CarData = CarData %>% filter(MSRP < 1000000)
 # Convert factor variables
@@
 ######### Try Something else to split data ############
-# First 842 are Models that only have 5 observations or less
+# First 836 are Models that only have 5 observations or less - deleted 6 card over 1 million dollars
 ModelList = CarData%>% group_by(Model) %>% mutate(ModelCount = n()) %>% arrange(ModelCount)
@@
 ModelList = ModelList[,-27]
-Not.Able.To.Split = ModelList[(1:842),]
-Able.To.Split = ModelList[-(1:842),]
+Not.Able.To.Split = ModelList[(1:836),]
+Able.To.Split = ModelList[-(1:836),]
 
@@
 train = bind_rows(train,Not.Able.To.Split)
+
 dim(train)
@@
 library(leaps)
-reg.fwd=regsubsets(MSRP~.,data=train,method="forward",nvmax=966)
+reg.fwd=regsubsets(MSRP~.,data=train,method="forward",nvmax=963)
 par(mfrow=c(1,3))
 bics<-summary(reg.fwd)$bic
-plot(1:967,bics,type="l",ylab="BIC",xlab="# of predictors")
+plot(1:964,bics,type="l",ylab="BIC",xlab="# of predictors")
 index<-which(bics==min(bics))
@@
 adjr2<-summary(reg.fwd)$adjr2
-plot(1:967,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
+plot(1:964,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
 index<-which(adjr2==max(adjr2))
@@
 rss<-summary(reg.fwd)$rss
-plot(1:967,rss,type="l",ylab="train RSS",xlab="# of predictors")
+plot(1:964,rss,type="l",ylab="train RSS",xlab="# of predictors")
 index<-which(rss==min(rss))
@@
 which(bics==min(bics))
-coef(reg.fwd,305)
+coef(reg.fwd,349)
 # By looking at this model, it has selected following predictors:
-# Make+Model+EngineFuelType+VehicleSize+VehicleStyle+Crossover+HighPerformance+TransmissionType+Exotic+Hatchback
-forward.regression = lm(MSRP~Make+Model+EngineFuelType+VehicleSize+VehicleStyle+Crossover+HighPerformance+TransmissionType+Exotic+Hatchback,data=train)
+# Make+Model+Year+EngineFuelType+NumberOfDoors+VehicleSize+VehicleStyle+CityMPG+Luxury+TransmissionType+Exotic+Hatchback
+forward.regression = lm(MSRP~Make+Model+Year+EngineFuelType+NumberOfDoors+VehicleSize+VehicleStyle+CityMPG+Luxury+TransmissionType+Exotic+Hatchback,data=train)
 summary(forward.regression)
@@
 ols_plot_resid_stand(forward.regression)
-# Cook's D plot looks okay - Much better than Log Transformed Plot
-#Observation 3748,9152,9153 seem to have problem on the residual plot
-summary(train[3748,]) # Ford Focus ... Not sure if we can delete this...
-summary(train[9152,]) # Bugatti ... Delete?
-summary(train[9153,]) # Bugatti ... Delete?
-summary(train[1043,]) # Volvo ... Not seem to cause too much trouble on Cook's D
+# Cook's D plot looks okay 
+# Residual plot... ???
 
@@
 ########################################
-reg.bwd=regsubsets(MSRP~.,data=train,method="backward",nvmax=966)
+reg.bwd=regsubsets(MSRP~.,data=train,method="backward",nvmax=963)
 par(mfrow=c(1,3))
 bics<-summary(reg.bwd)$bic
-plot(1:967,bics,type="l",ylab="BIC",xlab="# of predictors")
+plot(1:964,bics,type="l",ylab="BIC",xlab="# of predictors")
 index<-which(bics==min(bics))
@@
 adjr2<-summary(reg.bwd)$adjr2
-plot(1:967,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
+plot(1:964,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
 index<-which(adjr2==max(adjr2))
@@
 rss<-summary(reg.bwd)$rss
-plot(1:967,rss,type="l",ylab="train RSS",xlab="# of predictors")
+plot(1:964,rss,type="l",ylab="train RSS",xlab="# of predictors")
 index<-which(rss==min(rss))
@@
 which(bics==min(bics))
-coef(reg.bwd,248)
+coef(reg.bwd,410)
 # Predictor for backward elimination method based on lowest BIC value
-# Make+Model+EngineFuelType+VehicleSize+Crossover+HighPerformance+Hybrid+VehicleStyle+Exotic
-backward.regression = lm(MSRP~Make+Model+EngineFuelType+VehicleSize+Crossover+HighPerformance+Hybrid+VehicleStyle+Exotic,data=train)
+# Make+Model+EngineFuelType+NumberOfDoors+VehicleSize+VehicleStyle+CityMPG+Luxury+FactoryTuner+FlexFuel+Exotic+Hatchback
+backward.regression = lm(MSRP~Make+Model+EngineFuelType+NumberOfDoors+VehicleSize+VehicleStyle+CityMPG+Luxury+FactoryTuner+FlexFuel+Exotic+Hatchback,data=train)
 summary(backward.regression)
@@
 xtest<-model.matrix(MSRP~.,test)[,-1]
-ytest<-test$MSRP #36461071
+ytest<-test$MSRP #36302210
 
@@
 testMSE_Forward<-mean((test.result-forward.pred)^2)
-testMSE_Forward #46312425
+testMSE_Forward #46054176
 
@@
 testMSE_Backward<-mean((test.result-backward.pred)^2)
-testMSE_Backward #48317165
+testMSE_Backward #44935730
 

### 2021-01-26 MSDS6372Project1.Rmd
update lasso coefficients

@@
 # Look at the coefficient for best lambda
+# Make+Model+Year+EngineFuelType+EngineHP+EngineCylinders+TransmissionType+DrivenWheels+NumberOfDoors+VehicleSize+VehicleStyle+HighwayMPG+CityMPG+Popularity+Crossover+Diesel+Exotic+Luxury+HighPerformance+FactoryTurner+Performance+FlexFuel+Hatchback+Hybrid+Yes+N_A
+# Lasso turned off some predictors in Make, Model,VehicleStyle... In order to check assumptions, we need to use lm model with all the predictors
+# Although this won't imitate the Lasso model perfectly though.
 coefficients = coef(cv.out, cv.out$lambda.min)

### 2021-01-26 MSDS6372Project1.Rmd
clean up code restart objective 1

@@
 
-#### Objecive 1
-We are trying to analyse the importance of the "Popularity" variable.
-While the details of this variable is vague, it was created from social media, and the "higher ups" are curious how much general popularity can play a role in the retail price of a vehicle.
-
-```{r}
-library(tidyverse)
-library(ggplot2)
-library(GGally)
-# You need to adjust the work environment to run this code
-setwd("/Users/mingyang/Desktop/SMU/Applied Statistics/MSDS6372Project1")
-# Directly import cleaned dataset and start from here
-CarData = read.csv("CleanedCarData.csv")
-# Convert factor variables
-variables.to.factor = c("Make","Model","EngineFuelType","TransmissionType","DrivenWheels","MarketCategory","VehicleSize","VehicleStyle","Crossover","Diesel","Exotic","Luxury","HighPerformance","FactoryTuner","Performance","FlexFuel","Hatchback","Hybrid","N_A")
-CarData[variables.to.factor] = lapply(CarData[variables.to.factor],factor)
-summary(CarData)
-# Plot agsint numeric variables
-CarData %>% select(MSRP,Year,EngineHP,EngineCylinders,NumberOfDoors,HighwayMPG,CityMPG,Popularity) %>% ggpairs()
-# Try Log transformation on MSRP
-CarData$LogMSRP = log(CarData$MSRP)
-# Recheck the logMSRP against numeric variables
-# LogMSRP seem to have improved linear relationship between Year and EngineHP correlations
-CarData %>% select(LogMSRP,Year,EngineHP,EngineCylinders,NumberOfDoors,HighwayMPG,CityMPG,Popularity) %>% ggpairs()
-# Look at Popularity
-boxplot(CarData$Popularity)
-CarData %>% select(LogMSRP,Popularity)%>% ggpairs()
-#Try LogPopularity
-CarData$LogPopularity = log(CarData$Popularity)
-CarData %>% select(LogMSRP,LogPopularity)%>% ggpairs()
-# Look at original MSRP vs Popularity
-CarData %>% select(MSRP,Popularity)%>% ggpairs()
-#LogMSRP and LogPopularity seem to have improved the linearity relationship between the two
-# Look at histogram of LogMSRP
-CarData %>% ggplot(aes(x=LogMSRP))+geom_histogram(bins=100)
-# Drop MSRP and Popularity from the original Dataset
-CarData<-CarData[,-c(15,16)]
-# Drop MarketCategory - 72 levels are not useful for prediction
-CarData <-CarData[,-10]
-str(CarData)
-#Split data into 80% Train 10% Test 10% Validation set
-set.seed(111)
-train.test.index<-sample(1:dim(CarData)[1],10723,replace=F)
-train.test = CarData[train.test.index,]
-validate = CarData[-train.test.index,]
-set.seed(112)
-train.index = sample(1:dim(train.test)[1],9531,replace=F)
-train = train.test[train.index,]
-test = train.test[-train.index,]
-
-# Use Forward variable selection method to select possible variables as predictors
-library(leaps)
-reg.fwd=regsubsets(LogMSRP~.,data=train,method="forward",nvmax=955)
-par(mfrow=c(1,3))
-bics<-summary(reg.fwd)$bic
-plot(1:956,bics,type="l",ylab="BIC",xlab="# of predictors")
-index<-which(bics==min(bics))
-points(index,bics[index],col="red",pch=10)
-
-adjr2<-summary(reg.fwd)$adjr2
-plot(1:956,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
-index<-which(adjr2==max(adjr2))
-points(index,adjr2[index],col="red",pch=10)
-
-rss<-summary(reg.fwd)$rss
-plot(1:956,rss,type="l",ylab="train RSS",xlab="# of predictors")
-index<-which(rss==min(rss))
-points(index,rss[index],col="red",pch=10)
-
-coef(reg.fwd,676)
-# By using Forward method, 676 predictors seem to generate the smallest BIC number
-# This model has selected predictors below:
-# Make+Model+EngineFuelType+TransmissionType+VehicleSize+VehicleStyle+HighwayMPG+CityMPG+Luxury+FactoryTuner+Performance+FlexFuel+Hybrid+N_A+Exotic+Hatchback+LogPopularity
-forward.regression = lm(LogMSRP~Make+Model+EngineFuelType+TransmissionType+VehicleSize+VehicleStyle+HighwayMPG+CityMPG+Luxury+FactoryTuner+Performance+FlexFuel+Hybrid+N_A+Exotic+Hatchback+LogPopularity,data=train)
-summary(forward.regression)
-par(mfrow=c(2,2))
-plot(forward.regression)
-library(olsrr)
-ols_plot_cooksd_bar(forward.regression)
-ols_plot_resid_stand(forward.regression)
-# After observing the Cook's D, there is observation 4659, 9435 has extremely large Cook's D
-# Observation 1523 also has large Cook's D
-train[4659,] # Toyota Yaris iA
-exp(train[4659,25]) 
-train%>% filter(Make=="Toyota") %>% select(Make,Model,Year,LogMSRP)
-# Can't really make sense to eliminate observation 4659
-
-train[9435,] # Ferrari Superamerica 2005
-exp(train[9435,25]) # Exotic Car
-
-#Look at other Exotic Cars
-train%>%filter(Exotic=="Yes")%>% select(Make,Model,Year,LogMSRP,LogPopularity)
-# Don't really make sense to delete observation 4659 either
-
-# Since there are so many categorical variables, and log transformation improvement on Year and EngineHP didn't make it to the final model
-# Next section we will try reimport the dataset without transformation
-```
-
-#### Objective 1 continue without Log transformation - since there are so many categorical variables
-
+# Start Here... Load in cleanedData, delete over 1 million dollar cars
+## split dataset into train/test/validate sets with set seed number
+### Models that only have 5 observations will all end up in train set
 ```{r}
@@
 str(CarData)
+```
+
+#### Objecive 1
+We are trying to analyse the importance of the "Popularity" variable.
+While the details of this variable is vague, it was created from social media, and the "higher ups" are curious how much general popularity can play a role in the retail price of a vehicle.
+
+This model need to be highly interpretable.
+We need to provide interpretation of the regression coefficients of the final model including hypothesis testing, interpretation of regression coefficients, and confidence intervals.
+
+```{r}
+
+```
+
+#### Objective 2 start without Log transformation - since there are so many categorical variables
+
+```{r}
 # Use Forward variable selection method to select possible variables as predictors

### 2021-01-26 MSDS6372Project1.Rmd
change tibble to data frame
add three models for Objective 1

@@
 dim(train)
-
+train = as.data.frame(train)
+test = as.data.frame(test)
+validate = as.data.frame(validate)
 str(train)
@@
 ```{r}
+# Explore numerical variables
+train %>% select(MSRP,Year,EngineHP,EngineCylinders,NumberOfDoors,HighwayMPG,CityMPG,Popularity)%>%ggpairs()
+# Try Log transformation on MSRP see if it will improve correlations
+train$logMSRP = log(train$MSRP)
+train %>% select(logMSRP,Year,EngineHP,EngineCylinders,NumberOfDoors,HighwayMPG,CityMPG,Popularity)%>%ggpairs()
+# It looks like logMSRP has increased correlation with Year but decreased correlation with EngineHP, EngineCylinders
+# It's best to run LASSO and choose predictors on both with transformation and without Log transformation in order to identify which
+# Model performs better
+train = train[,-27]
+#str(train)
+library(glmnet)
+# Since we want a easy to interperate model, we don't want to use predictor Make and Model that has too many levels in the model
+train = train[,-(1:2)]
+str(train)
+#Dummy code categorical predictor variables
+x <- model.matrix(MSRP~.,train)[,-1]
+y <- log(train$MSRP)
+set.seed(123) # Need a seed number to keep same results for LASSO
+cv.out=cv.glmnet(x,y,alpha=1) #alpha=1 performs LASSO
+par(mfrow=c(1,1))
+plot(cv.out)
+# Look at coefficients
+coef(cv.out, cv.out$lambda.min)
+# Seems like LASSO has selected all predictors, only few levels within EngineFuel, TransmissionType, VehicleStyle got turned off
+# Build model with predictors LASSO selected
+log.model1 = lm(log(MSRP)~.,data=train)
+par(mfrow=c(2,2))
+plot(log.model1)
+library(olsrr)
+ols_plot_cooksd_bar(log.model1)
+ols_plot_resid_stand(log.model1)
+# Residual looks okay~
+# Now try MSRP without log transformation
 
+#Dummy code categorical predictor variables
+x1 <- model.matrix(MSRP~.,train)[,-1]
+y1 <- train$MSRP
+set.seed(123) # Need a seed number to keep same results for LASSO
+cv.out1=cv.glmnet(x1,y1,alpha=1) #alpha=1 performs LASSO
+par(mfrow=c(1,1))
+plot(cv.out1)
+# Look at coefficients
+coef(cv.out1, cv.out1$lambda.min)
+# Again, all predictors are selected, only few levels are turned off
+regular.model1 = lm(MSRP~.,data=train)
+par(mfrow=c(2,2))
+plot(regular.model1)
+ols_plot_cooksd_bar(regular.model1)
+ols_plot_resid_stand(regular.model1)
+
+#############################################################################
+# By comparison Constant variance looks a lot better with the log(MSRP)...###
+# Since the assumption works best with this, next check with stepwise method to see what predictors will be selected
+stepwise.model1 =lm(log(MSRP)~.,data=train)
+ols_step_both_aic(stepwise.model1)
+# By Using StepwiseAIC variable selection method, the following predictors are selected
+# Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+HighwayMPG+Hybrid+TransmissionType+EngineCylinders+Luxury+VehicleSize+FlexFuel+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+CityMPG+Popularity+N_A+Diesel
+log.model2.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+HighwayMPG+Hybrid+TransmissionType+EngineCylinders+Luxury+VehicleSize+FlexFuel+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+CityMPG+Popularity+N_A+Diesel,data=train)
+par(mfrow=c(2,2))
+plot(log.model2.step)
+ols_plot_cooksd_bar(log.model2.step)
+ols_plot_resid_stand(log.model2.step)
+
+#compare test ASE results between stepwise model and 
+test = test[,-(1:2)]
+testMSRP = test$MSRP # GET real result 
+test = test[,-13] # Delete MSRP from test 
+# Calculate stepwise test ASE first
+pred.test.step = predict(log.model2.step,test)
+# exp values to return to original value
+pred.test.step = exp(pred.test.step)
+test_ASE_STEP = mean((testMSRP-pred.test.step)^2)
+test_ASE_STEP #497620223
+
+#Calculate LASSO test ASE
+pred.test.lasso = predict(log.model1,test)
+pred.test.lasso = exp(pred.test.lasso)
+test_ASE_LASSO = mean((testMSRP-pred.test.lasso)^2)
+test_ASE_LASSO #499722642
+
+# By comparison, it seems LASSO performs a bit worse than StepwiseAIC selected predictors...
 ```
 
+# Note: Rerun the code chunk for clean train/test/validate datasets - altered in Objective 1
 #### Objective 2 start without Log transformation - since there are so many categorical variables

### 2021-01-27 MSDS6372Project1.Rmd
adjusting models for objective 1 with VIF and Summary for selecting the best predictors for simple model.

--- 2021-01-26 MSDS6372Project1.Rmd
@@
 ```{r}
+# Explore numerical variables
+train %>% select(MSRP,Year,EngineHP,EngineCylinders,NumberOfDoors,HighwayMPG,CityMPG,Popularity)%>%ggpairs()
+# Try Log transformation on MSRP see if it will improve correlations
+train$logMSRP = log(train$MSRP)
+train %>% select(logMSRP,Year,EngineHP,EngineCylinders,NumberOfDoors,HighwayMPG,CityMPG,Popularity)%>%ggpairs()
+# It looks like logMSRP has increased correlation with Year but decreased correlation with EngineHP, EngineCylinders
+# It's best to run LASSO and choose predictors on both with transformation and without Log transformation in order to identify which
+# Model performs better
+train = train[,-27]
+#str(train)
+library(glmnet)
+library(car)
+# Since we want a easy to interperate model, we don't want to use predictor Make and Model that has too many levels in the model
+train = train[,-(1:2)]
+str(train)
+#Dummy code categorical predictor variables
+x <- model.matrix(MSRP~.,train)[,-1]
+y <- log(train$MSRP)
+set.seed(123) # Need a seed number to keep same results for LASSO
+cv.out=cv.glmnet(x,y,alpha=1) #alpha=1 performs LASSO
+par(mfrow=c(1,1))
+plot(cv.out)
+# Look at coefficients
+coef(cv.out, cv.out$lambda.min)
+# Seems like LASSO has selected all predictors, only few levels within EngineFuel, TransmissionType, VehicleStyle got turned off
+# Build model with predictors LASSO selected
+log.model1 = lm(log(MSRP)~.,data=train)
+# Check VIF
+summary(log.model1)
+# TransmissionType and Hatchback provides NA on summary statistic, try remove them as predictors
+log.model2 = lm(log(MSRP)~.- TransmissionType - Hatchback,data=train)
+summary(log.model2)
+# Look at VIF
+vif(log.model2)[,3]^2
+# CityMPG has really high VIF remove it from model
+log.model3 = lm(log(MSRP)~.- TransmissionType - Hatchback - CityMPG,data=train)
+# Look at VIF
+vif(log.model3)[,3]^2
+# NumberOfDoors has really high VIF remove from model
+log.model4 = lm(log(MSRP)~.- TransmissionType - Hatchback - CityMPG - NumberOfDoors,data=train)
+# Look at VIF
+vif(log.model4)[,3]^2
+# HighwayMPG still have over 10 VIF, remove from model
+log.model5 = lm(log(MSRP)~.- TransmissionType - Hatchback - CityMPG - NumberOfDoors - HighwayMPG,data=train)
+vif(log.model5)[,3]^2
+# Diesel is not significant, HighPerformance is not significant, Performance is not significant Remove from simple model
+log.model6 = lm(log(MSRP)~.- TransmissionType - Hatchback - CityMPG - NumberOfDoors - HighwayMPG - Diesel - HighPerformance - Performance,data=train)
+vif(log.model6)[,3]^2
+# N_A is still not significant, remove from model
+log.model7 = lm(log(MSRP)~.- TransmissionType - Hatchback - CityMPG - NumberOfDoors - HighwayMPG - Diesel - HighPerformance - Performance - N_A,data=train)
+vif(log.model7)[,3]^2
+summary(log.model7)
+# Model 7 is a highly interpretable model with decent R-squared.
 
+par(mfrow=c(2,2))
+plot(log.model7)
+library(olsrr)
+ols_plot_cooksd_bar(log.model1)
+ols_plot_resid_stand(log.model1)
+# Residual looks okay~
+# Now try MSRP without log transformation
+
+#Dummy code categorical predictor variables
+x1 <- model.matrix(MSRP~.,train)[,-1]
+y1 <- train$MSRP
+set.seed(123) # Need a seed number to keep same results for LASSO
+cv.out1=cv.glmnet(x1,y1,alpha=1) #alpha=1 performs LASSO
+par(mfrow=c(1,1))
+plot(cv.out1)
+# Look at coefficients
+coef(cv.out1, cv.out1$lambda.min)
+# Again, all predictors are selected, only few levels are turned off
+regular.model1 = lm(MSRP~.,data=train)
+summary(regular.model1)
+# Delete Hatchback & TransmissionType... NA on summary table
+regular.model2 = lm(MSRP~.-Hatchback-TransmissionType,data=train)
+vif(regular.model2)[,3]^2
+# remove CityMPG - highest VIF, larger than 29
+regular.model3 = lm(MSRP~.-Hatchback-TransmissionType-CityMPG,data=train)
+vif(regular.model3)[,3]^2
+# remove NumberOfDoors
+regular.model4 = lm(MSRP~.-Hatchback-TransmissionType-CityMPG-NumberOfDoors,data=train)
+vif(regular.model4)[,3]^2
+# HighwayMPG still over 10
+regular.model5 = lm(MSRP~.-Hatchback-TransmissionType-CityMPG-NumberOfDoors-HighwayMPG,data=train)
+vif(regular.model5)[,3]^2
+summary(regular.model5)
+# FlexFuel has the highest P value in terms of significance - delete
+regular.model6 = lm(MSRP~.-Hatchback-TransmissionType-CityMPG-NumberOfDoors-HighwayMPG-FlexFuel,data=train)
+vif(regular.model6)[,3]^2
+summary(regular.model6)
+#Diesel is high in p-value delete
+regular.model7 = lm(MSRP~.-Hatchback-TransmissionType-CityMPG-NumberOfDoors-HighwayMPG-FlexFuel-Diesel,data=train)
+vif(regular.model7)[,3]^2
+summary(regular.model7)
+par(mfrow=c(2,2))
+plot(regular.model1) # Residual plot doesn't look as well as log(MSRP) residual so tranformation is necessary.
+ols_plot_cooksd_bar(regular.model1)
+ols_plot_resid_stand(regular.model1)
+
+#############################################################################
+# By comparison Constant variance looks a lot better with the log(MSRP)...###
+# Since the assumption works best with this, next check with stepwise method to see what predictors will be selected
+stepwise.model1 =lm(log(MSRP)~.,data=train)
+ols_step_both_aic(stepwise.model1)
+# By Using StepwiseAIC variable selection method, the following predictors are selected
+# Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+HighwayMPG+Hybrid+TransmissionType+EngineCylinders+Luxury+VehicleSize+FlexFuel+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+CityMPG+Popularity+N_A+Diesel
+log.model2.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+HighwayMPG+Hybrid+TransmissionType+EngineCylinders+Luxury+VehicleSize+FlexFuel+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+CityMPG+Popularity+N_A+Diesel,data=train)
+summary(log.model2.step)
+# TransmissionType has NA need to be deleted
+log.model3.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+HighwayMPG+Hybrid+EngineCylinders+Luxury+VehicleSize+FlexFuel+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+CityMPG+Popularity+N_A+Diesel,data=train)
+summary(log.model3.step)
+vif(log.model3.step)[,3]^2
+# CityMPG has the highest VIF - delete
+log.model4.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+HighwayMPG+Hybrid+EngineCylinders+Luxury+VehicleSize+FlexFuel+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+Popularity+N_A+Diesel,data=train)
+vif(log.model4.step)[,3]^2
+# HighwayMPG still over 10, delete
+log.model5.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+Hybrid+EngineCylinders+Luxury+VehicleSize+FlexFuel+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+Popularity+N_A+Diesel,data=train)
+vif(log.model5.step)[,3]^2
+summary(log.model5.step)
+#FlexFuel has the highest insignificant P-value delete
+log.model6.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+Hybrid+EngineCylinders+Luxury+VehicleSize+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+Popularity+N_A+Diesel,data=train)
+summary(log.model6.step)
+#Diesel has the highest insignificant P-value delete
+log.model7.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+Hybrid+EngineCylinders+Luxury+VehicleSize+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+Popularity+N_A,data=train)
+summary(log.model7.step)
+#N_A has the highest insignificant P-value delete
+log.model8.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+Hybrid+EngineCylinders+Luxury+VehicleSize+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+Popularity,data=train)
+summary(log.model8.step)
+
+par(mfrow=c(2,2))
+plot(log.model8.step)
+ols_plot_cooksd_bar(log.model8.step)
+ols_plot_resid_stand(log.model8.step)
+
+# LASSO and STEPWISE came to slight different Model in terms of predictor and r-squared value.
+# LASSO r-squared 0.8426, adjusted r-squared 0.8419
+# Stepwise r-squared 0.843, adjusted r-squared 0.8424
+#compare test ASE results between stepwise model and LASSO
+test = test[,-(1:2)]
+testMSRP = test$MSRP # GET real result 
+test = test[,-13] # Delete MSRP from test 
+# Calculate stepwise test ASE first
+pred.test.step = predict(log.model8.step,test)
+# exp values to return to original value
+pred.test.step = exp(pred.test.step)
+test_ASE_STEP = mean((testMSRP-pred.test.step)^2)
+test_ASE_STEP #560738819
+
+#Calculate LASSO test ASE
+pred.test.lasso = predict(log.model7,test)
+pred.test.lasso = exp(pred.test.lasso)
+test_ASE_LASSO = mean((testMSRP-pred.test.lasso)^2)
+test_ASE_LASSO #571497373
+
+# By comparison, Stepwise model workflow provided the best model in terms of r-squared, adjusted r-squared and test ASE.
+# Move forward with the analysis with log.model8.step
 ```
 
+# Note: Rerun the code chunk for clean train/test/validate datasets - altered in Objective 1
 #### Objective 2 start without Log transformation - since there are so many categorical variables

### 2021-01-31 MSDS6372Project1.Rmd
Fixing Flex Fuel section

@@
 CarData$Performance = ifelse(grepl("Performance",CarData$Market.Category)&(!grepl("High-Performance",CarData$Market.Category)),"Yes","No")
+# Flex Fuel column if EngineFuelType has flex-fuel
+CarData$`Flex Fuel` = ifelse(grepl("flex-fuel",CarData$Engine.Fuel.Type)&CarData$`Flex Fuel`=="No","Yes",CarData$`Flex Fuel`)
 
@@
 # Look at summary of data
-summary(CarData$VehicleStyle)
 str(CarData)
@@
 vif(log.model5)[,3]^2
+summary(log.model5)
 # Diesel is not significant, HighPerformance is not significant, Performance is not significant Remove from simple model
@@
 vif(log.model6)[,3]^2
+summary(log.model6)
 # N_A is still not significant, remove from model

### 2021-01-31 MSDS6372Project1.Rmd
update objective 1

@@
 summary(log.model5)
-# Diesel is not significant, HighPerformance is not significant, Performance is not significant Remove from simple model
-log.model6 = lm(log(MSRP)~.- TransmissionType - Hatchback - CityMPG - NumberOfDoors - HighwayMPG - Diesel - HighPerformance - Performance,data=train)
+# HighPerformance has the least statistical significance - delete
+log.model6 = lm(log(MSRP)~.- TransmissionType - Hatchback - CityMPG - NumberOfDoors - HighwayMPG - HighPerformance, data=train)
 vif(log.model6)[,3]^2
 summary(log.model6)
-# N_A is still not significant, remove from model
-log.model7 = lm(log(MSRP)~.- TransmissionType - Hatchback - CityMPG - NumberOfDoors - HighwayMPG - Diesel - HighPerformance - Performance - N_A,data=train)
+# Diesel has the least statistical significance - delete
+log.model6 = lm(log(MSRP)~.- TransmissionType - Hatchback - CityMPG - NumberOfDoors - HighwayMPG - HighPerformance - Diesel, data=train)
+vif(log.model6)[,3]^2
+summary(log.model6)
+# N_A has the least statistical significance - delete
+log.model6 = lm(log(MSRP)~.- TransmissionType - Hatchback - CityMPG - NumberOfDoors - HighwayMPG - HighPerformance - Diesel - N_A, data=train)
+vif(log.model6)[,3]^2
+summary(log.model6)
+# Performance is still not significant, remove from model
+log.model7 = lm(log(MSRP)~.- TransmissionType - Hatchback - CityMPG - NumberOfDoors - HighwayMPG - HighPerformance - Diesel - N_A - Performance, data=train)
 vif(log.model7)[,3]^2
@@
 # By Using StepwiseAIC variable selection method, the following predictors are selected
-# Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+HighwayMPG+Hybrid+TransmissionType+EngineCylinders+Luxury+VehicleSize+FlexFuel+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+CityMPG+Popularity+N_A+Diesel
-log.model2.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+HighwayMPG+Hybrid+TransmissionType+EngineCylinders+Luxury+VehicleSize+FlexFuel+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+CityMPG+Popularity+N_A+Diesel,data=train)
+# Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+HighwayMPG+Hybrid+TransmissionType+EngineCylinders+Luxury+VehicleSize+DrivenWheels+FlexFuel+NumberOfDoors+Crossover+FactoryTuner+Popularity+CityMPG+N_A
+log.model2.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+HighwayMPG+Hybrid+TransmissionType+EngineCylinders+Luxury+VehicleSize+DrivenWheels+FlexFuel+NumberOfDoors+Crossover+FactoryTuner+Popularity+CityMPG+N_A,data=train)
 summary(log.model2.step)
 # TransmissionType has NA need to be deleted
-log.model3.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+HighwayMPG+Hybrid+EngineCylinders+Luxury+VehicleSize+FlexFuel+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+CityMPG+Popularity+N_A+Diesel,data=train)
+log.model3.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+HighwayMPG+Hybrid+EngineCylinders+Luxury+VehicleSize+DrivenWheels+FlexFuel+NumberOfDoors+Crossover+FactoryTuner+Popularity+CityMPG+N_A,data=train)
 summary(log.model3.step)
@@
 # CityMPG has the highest VIF - delete
-log.model4.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+HighwayMPG+Hybrid+EngineCylinders+Luxury+VehicleSize+FlexFuel+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+Popularity+N_A+Diesel,data=train)
+log.model4.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+HighwayMPG+Hybrid+EngineCylinders+Luxury+VehicleSize+DrivenWheels+FlexFuel+NumberOfDoors+Crossover+FactoryTuner+Popularity+N_A,data=train)
 vif(log.model4.step)[,3]^2
-# HighwayMPG still over 10, delete
-log.model5.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+Hybrid+EngineCylinders+Luxury+VehicleSize+FlexFuel+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+Popularity+N_A+Diesel,data=train)
+# NumberOfDoors highest VIF - delete
+log.model5.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+HighwayMPG+Hybrid+EngineCylinders+Luxury+VehicleSize+DrivenWheels+FlexFuel+Crossover+FactoryTuner+Popularity+N_A,data=train)
+vif(log.model5.step)[,3]^2
+# HighwayMPG still over 10 - delete
+log.model5.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+Hybrid+EngineCylinders+Luxury+VehicleSize+DrivenWheels+FlexFuel+Crossover+FactoryTuner+Popularity+N_A,data=train)
 vif(log.model5.step)[,3]^2
 summary(log.model5.step)
-#FlexFuel has the highest insignificant P-value delete
-log.model6.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+Hybrid+EngineCylinders+Luxury+VehicleSize+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+Popularity+N_A+Diesel,data=train)
-summary(log.model6.step)
-#Diesel has the highest insignificant P-value delete
-log.model7.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+Hybrid+EngineCylinders+Luxury+VehicleSize+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+Popularity+N_A,data=train)
-summary(log.model7.step)
 #N_A has the highest insignificant P-value delete
-log.model8.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+Hybrid+EngineCylinders+Luxury+VehicleSize+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+Popularity,data=train)
+log.model8.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+Hybrid+EngineCylinders+Luxury+VehicleSize+DrivenWheels+FlexFuel+Crossover+FactoryTuner+Popularity,data=train)
 summary(log.model8.step)
@@
 test_ASE_STEP = mean((testMSRP-pred.test.step)^2)
-test_ASE_STEP #560738819
+test_ASE_STEP #571362858
 
@@
 test_ASE_LASSO = mean((testMSRP-pred.test.lasso)^2)
-test_ASE_LASSO #571497373
+test_ASE_LASSO #571362858
 

### 2021-02-01 MSDS6372Project1.Rmd
objective 2 LogMSRP vs Degree2 EngineHP converting

@@
 lasso.model = glmnet(x,y,alpha = 1,lambda = cv.out$lambda.min)
-# Check assumptions
-
 #Dummy code test set to see performance compare to other methods
@@
 testMSE_LASSO
+################
+# In order to look at assumption, need to refit the model based on LASSO selected variables...
+# Check coefficients and see what got turn off:
+# - Crossover (only one turned off, others are within factor levels that are significant)
+################
+lasso.lm = lm(MSRP~.-Crossover,data=train)
+summary(lasso.lm)
+# Check Assumptions
+par(mfrow=c(2,2))
+plot(lasso.lm)
+library(olsrr)
+ols_plot_cooksd_bar(lasso.lm)
+ols_plot_resid_stand(lasso.lm)
+
+##################################################################################
+# Try Log(MSRP vs cubic Year & Second degree EngineHP variable selection #########
+##################################################################################
+
+# Above are findings during EDA... connections to logMSRP
+engineHP.lm = lm(log(MSRP)~EngineHP,data=train)
+ols_plot_resid_stand(engineHP.lm)
+# Looks like Second degree might help
+engineHP.lm2 = lm(log(MSRP)~EngineHP+I(EngineHP^2),data=train)
+ols_plot_resid_stand(engineHP.lm2)
+
+#Without transformation
+engineHP.lm = lm(MSRP~EngineHP,data=train)
+ols_plot_resid_stand(engineHP.lm)
+# Looks like Second degree might help, and MSRP looks better than log(MSRP)
+engineHP.lm2 = lm(MSRP~EngineHP+I(EngineHP^2),data=train)
+ols_plot_resid_stand(engineHP.lm2)
+
+# Try year
+year.lm = lm(log(MSRP)~Year,data=train)
+ols_plot_resid_stand(year.lm)
+# Try Second Degree Year
+year.lm2 = lm(log(MSRP)~Year+I(Year^2),data=train)
+ols_plot_resid_stand(year.lm2)
+# Try Third Degree Year
+year.lm3 = lm(log(MSRP)~Year+I(Year^2)+I(Year^3),data=train)
+ols_plot_resid_stand(year.lm3)
+#Try regular MSRP
+# Try year
+year.lm = lm(MSRP~Year,data=train)
+ols_plot_resid_stand(year.lm)
+# Try Second Degree Year
+year.lm2 = lm(MSRP~Year+I(Year^2),data=train)
+ols_plot_resid_stand(year.lm2)
+# Try Third Degree Year
+year.lm3 = lm(MSRP~Year+I(Year^2)+I(Year^3),data=train)
+ols_plot_resid_stand(year.lm3)
+############### Looks like LogMSRP vs year degree 3 helps
+############### Looks like LogMSRP vs EngineHP degree 2 helps
+# Reselect variables with LASSO
+#Dummy code categorical predictor variables
+x <- model.matrix(log(MSRP)~.+I(EngineHP^2)+I(Year^2)+I(Year^3),train)[,-1]
+y <- log(train$MSRP)
+set.seed(123) # Need a seed number to keep same results for LASSO
+cv.out2=cv.glmnet(x,y,alpha=1) #alpha=1 performs LASSO
+par(mfrow=c(1,1))
+plot(cv.out2)
+coef(cv.out2, cv.out2$lambda.min)
+# Year-Degree2 didn't get turned on neither did I(Year^3), I(EngineHP^2) did get turned on
+# Following also get turned off...
+# - Hatchback - Popularity 
+# Refit the model with predictors
+lasso.lm.log = lm(log(MSRP)~.+I(EngineHP^2)-Hatchback-Popularity,data=train)
+summary(lasso.lm.log) # R-Squared and Adjusted R-squared is slightly lower compared to regular MSRP using LASSO
+# Check Assumptions
+par(mfrow=c(2,2))
+plot(lasso.lm.log)
+library(olsrr)
+ols_plot_cooksd_bar(lasso.lm.log)
+ols_plot_resid_stand(lasso.lm.log)
+#Residual looks okay... check test ASE below
 
@@
 
+#LASSO test result
+lasso.pred2 = predict(lasso.lm,test)
+# See LASSO test ASE after refit
+testASE_LASSO_2 = mean((test.result-lasso.pred2)^2)
+testASE_LASSO_2 #35982124
+
+# LogMSRP with Second Degree EngineHP picked by LASSO test result
+lasso.log.pred1 = predict(lasso.lm.log,test)
+# Recover the value by exp
+lasso.log.pred1 = exp(lasso.log.pred1)
+testASE_LASSO_LOG = mean((test.result-lasso.log.pred1)^2)
+testASE_LASSO_LOG #24577120
+
 ```

### 2021-02-01 MSDS6372Project1.Rmd
testing

@@
 library(olsrr)
-ols_plot_cooksd_bar(log.model1)
-ols_plot_resid_stand(log.model1)
+ols_plot_cooksd_bar(log.model7)
+ols_plot_resid_stand(log.model7)
 # Residual looks okay~

### 2021-02-02 MSDS6372Project1.Rmd
upload tree regression and update test RMSE

@@
 test_ASE_STEP = mean((testMSRP-pred.test.step)^2)
-test_ASE_STEP #571362858
+sqrt(test_ASE_STEP) #test RMSE = 23903.2
 
@@
 test_ASE_LASSO = mean((testMSRP-pred.test.lasso)^2)
-test_ASE_LASSO #571362858
+sqrt(test_ASE_LASSO) ##test RMSE = 23903.2
 
@@
 testASE_LASSO_LOG = mean((test.result-lasso.log.pred1)^2)
-testASE_LASSO_LOG #24577120
+sqrt(testASE_LASSO_LOG) #test RMSE = 4957.532
 
 ```
+
+# Note: Rerun the code chunk for clean train/test/validate datasets
+#### Objective 2 KNN Regression
+KNN Regression is a non-parametric machine learning approach, it first identifies the K training observations that are closest to X0, it then estimates the response using the average of all the training response identified.
+The optimal value for K will depend on the bias-variance trade off.
+A small K provides the most flexible fit, which will have low bias but high variance.
+This variance is due to the fact that the prediction in a given region is entirely dependent on just one observation.
+In contrast, larger values of K provide a smoother and less variable fit; the prediction in a region is an average of several points, and so changing one observation has smaller effect.
+However, the smoothing may cause bias by masking some of the structure in the response.
+-- (ISLR Seventh Edition) Note: KNN regression can only handles numerical variables.
+
+```{r}
+library(caret)
+############################################################
+#testing KNN regression - only handles numeric variables####
+############################################################
+
+#getting training response
+trainY = train$MSRP
+#getting training predictors
+trainX = train[,-c(1,2,4,7,8,10,11,15,16,17,18,19,20,21,22,23,24,25,26)]
+# Doing the same for test
+testY = test$MSRP
+testX = test[,-c(1,2,4,7,8,10,11,15,16,17,18,19,20,21,22,23,24,25,26)]
+
+fit = knnreg(trainX,trainY,k=3)
+plot(testY,predict(fit,testX))
+#Try testing testASE
+knn.k3.pred = predict(fit,testX)
+testASE_KNN_k3 = mean((testY-knn.k3.pred)^2)
+sqrt(testASE_KNN_k3)
+
+############# explore optimal k using test ASE ###############
+iterations = 100
+testASE_List = c()
+for(j in 1:iterations){
+  fit = knnreg(trainX,trainY,k=j)
+  knn.pred = predict(fit,testX)
+  testASE_loop = mean((testY-knn.pred)^2)
+  testASE_List[j] = sqrt(testASE_loop)
+}
+par(mfrow=c(1,1))
+plot(1:iterations,testASE_List,type="l",xlab="# of k",ylab="test Root Mean Square Error")
+index<-which(testASE_List==min(testASE_List))
+index
+
+# seems k=2 gives the best result according to our test ASE
+fit = knnreg(trainX,trainY,k=2)
+plot(testY,predict(fit,testX))
+#Try testing testASE
+knn.k2.pred = predict(fit,testX)
+testASE_KNN_k2 = mean((testY-knn.k2.pred)^2)
+sqrt(testASE_KNN_k2)
+
+```

### 2021-02-02 MSDS6372Project1.Rmd
update test RMSE for regression tree

@@
 testMSE_Forward<-mean((test.result-forward.pred)^2)
-testMSE_Forward #46054176
+sqrt(testMSE_Forward) #test RMSE = 6786.323
 
@@
 testMSE_Backward<-mean((test.result-backward.pred)^2)
-testMSE_Backward #44935730
+sqrt(testMSE_Backward) #test RMSE = 6702.666
 
@@
 testASE_LASSO_2 = mean((test.result-lasso.pred2)^2)
-testASE_LASSO_2 #35982124
+sqrt(testASE_LASSO_2) #test RMSE = 5998.51
 
@@
 testASE_KNN_k2 = mean((testY-knn.k2.pred)^2)
-sqrt(testASE_KNN_k2)
+sqrt(testASE_KNN_k2) # test RMSE = 6861.357
 

### 2021-02-02 MSDS6372Project1.Rmd
adding interaction to objective 2 regression

@@
 
+######################################################################################################################
+# Try Log(MSRP) Second degree EngineHP, adding interaction EngineFuelType*EngineCylinders variable selection #########
+######################################################################################################################
+
+tryplot = train%>% select(MSRP,EngineCylinders,EngineFuelType)
+tryplot$logMSRP = log(tryplot$MSRP)
+tryplot %>% ggplot(aes(x=EngineCylinders,y=logMSRP, color=EngineFuelType))+
+  geom_point(position="jitter")+ scale_color_viridis_d()
+#Try with LASSO variable selection
+x <- model.matrix(log(MSRP)~.+I(EngineHP^2)+EngineFuelType*EngineCylinders,train)[,-1]
+y <- log(train$MSRP)
+set.seed(123) # Need a seed number to keep same results for LASSO
+cv.out3=cv.glmnet(x,y,alpha=1) #alpha=1 performs LASSO
+par(mfrow=c(1,1))
+plot(cv.out3)
+coef(cv.out3, cv.out3$lambda.min)
+# - Hatchback - Popularity
+# Refit the model with predictors
+lasso.lm.log.sq.int = lm(log(MSRP)~.+I(EngineHP^2)-Hatchback-Popularity+EngineFuelType*EngineCylinders,data=train)
+summary(lasso.lm.log.sq.int) # R-Squared and Adjusted R-squared is slightly lower compared to regular MSRP using LASSO
+# Check Assumptions
+par(mfrow=c(2,2))
+plot(lasso.lm.log.sq.int)
+library(olsrr)
+ols_plot_cooksd_bar(lasso.lm.log)
+ols_plot_resid_stand(lasso.lm.log)
+
 #####################################################
@@
 sqrt(testASE_LASSO_LOG) #test RMSE = 4957.532
+
+# LogMSRP with Second Degree EngineHP and interaction EngineCylinders*EngineFuelType
+lasso.log.sq.int.pred = predict(lasso.lm.log.sq.int,test)
+# Recover the value by exp
+lasso.log.sq.int.pred = exp(lasso.log.sq.int.pred)
+testASE_LASSO_LOG_Inter = mean((test.result-lasso.log.sq.int.pred)^2)
+sqrt(testASE_LASSO_LOG_Inter) #test RMSE = 4990.447
 

### 2021-02-02 MSDS6372Project1.Rmd
add validation result

@@
 ```
+
+# Validation test between best models under each objective
+
+```{r}
+validate.result = validate$MSRP
+validate = validate[,-15]
+# Validate best model from objective 1 result
+objective1.predict = predict(log.model8.step,validate)
+# See forward test ASE
+validateMSE_Object1 <- mean((validate.result-objective1.predict)^2)
+sqrt(validateMSE_Object1) #test RMSE = 55452.96
+
+# Validate best model from objective 2 result
+objective2.reg.predict = predict(lasso.lm.log,validate)
+validateMSE_object2.reg.predict = mean((validate.result-objective2.reg.predict)^2)
+sqrt(validateMSE_object2.reg.predict) #test RMSE = 55452.94
+
+# Validate best model from KNN
+validateKNN = validate[,-c(1,2,4,7,8,10,11,15,16,17,18,19,20,21,22,23,24,25)]
+objective2.knn.predict = predict(fit,validateKNN)
+validateASE_KNN_k2 = mean((validate.result-objective2.knn.predict)^2)
+sqrt(validateASE_KNN_k2) # test RMSE = 5718.328
+
+```

### 2021-02-02 MSDS6372Project1.Rmd
add validation result 2

--- 2021-02-02 MSDS6372Project1.Rmd
@@
 ```
+
+# Validation test between best models under each objective
+
+```{r}
+validate.result = validate$MSRP
+validate = validate[,-15]
+# Validate best model from objective 1 result
+objective1.predict = predict(log.model7,validate)
+objective1.predict = exp(objective1.predict)
+# See forward test ASE
+validateMSE_Object1 <- mean((validate.result-objective1.predict)^2)
+sqrt(validateMSE_Object1) #test RMSE = 21896.2
+
+# Validate best model from objective 2 result
+objective2.reg.predict = predict(lasso.lm.log,validate)
+objective2.reg.predict = exp(objective2.reg.predict)
+validateMSE_object2.reg.predict = mean((validate.result-objective2.reg.predict)^2)
+sqrt(validateMSE_object2.reg.predict) #test RMSE = 5521.819
+
+# Validate best model from KNN
+validateKNN = validate[,-c(1,2,4,7,8,10,11,15,16,17,18,19,20,21,22,23,24,25)]
+objective2.knn.predict = predict(fit,validateKNN)
+validateASE_KNN_k2 = mean((validate.result-objective2.knn.predict)^2)
+sqrt(validateASE_KNN_k2) # test RMSE = 5718.328
+
+```

### 2021-02-02 MSDS6372Project1.Rmd
tried scaling method, commented out

@@
 trainX = train[,-c(1,2,4,7,8,10,11,15,16,17,18,19,20,21,22,23,24,25,26)]
+#try scaling variables
+#str(trainX)
+#for(i in 1: dim(trainX)[2]){
+#  trainX[,i] = scale(trainX[,i])
+#}
+# finished scaling
 # Doing the same for test
@@
 testX = test[,-c(1,2,4,7,8,10,11,15,16,17,18,19,20,21,22,23,24,25,26)]
+#str(testX)
+#for(i in 1: dim(testX)[2]){
+#  testX[,i] = scale(testX[,i])
+#}
 
